---
title: "Homework 6"
author: Jingyi Zhang
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(modelr)
library(p8105.datasets)
library(purrr)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))
          
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

```{r load_clean_data_1}
homicide_df =
  read_csv("./homicide_data/homicide-data.csv", na = c("", "NA", "Unknown")) %>% 
  mutate(
    city_state = str_c(city, state, sep = ", "),
    victim_age = as.numeric(victim_age),
    resolution = case_when(
      disposition == "Closed without arrest" ~ 0,
      disposition == "Open/No arrest"        ~ 0,
      disposition == "Closed by arrest"      ~ 1)
  ) %>% 
  filter(
    victim_race %in% c("White", "Black"),
    city_state != "Tulsa, AL") %>% 
  select(city_state, resolution, victim_age, victim_race, victim_sex)
```

<br />

### Start with one city

```{r}
baltimore_df =
  homicide_df %>% 
  filter(city_state == "Baltimore, MD")

# fit a generalized linear model
glm(resolution ~ victim_age + victim_race + victim_sex, 
    data = baltimore_df,
    family = binomial()) %>% 
  broom::tidy() %>% 
  mutate(
    OR = exp(estimate), # odds ratio
    CI_lower = exp(estimate - 1.96 * std.error),
    CI_upper = exp(estimate + 1.96 * std.error)
  ) %>% 
  select(term, OR, starts_with("CI")) %>% 
  knitr::kable(digits = 3)
```

<br />

### Try this across cities

```{r}
models_results_df =
  homicide_df %>% 
  nest(data = -city_state) %>% 
  mutate(
    models = 
      map(.x = data, ~glm(resolution ~ victim_age + victim_race + victim_sex, data = .x, family = binomial())),
    results = map(models, broom::tidy)
    ) %>% 
  select(city_state, results) %>% 
  unnest(results) %>% 
  mutate(
    OR = exp(estimate), # odds ratio
    CI_lower = exp(estimate - 1.96 * std.error),
    CI_upper = exp(estimate + 1.96 * std.error)
  ) %>% 
  select(city_state, term, OR, starts_with("CI"))
```

```{r}
models_results_df %>% 
  filter(term == "victim_sexMale") %>% 
  mutate(city_state = fct_reorder(city_state, OR)) %>% 
  ggplot(aes(x = city_state, y = OR)) +
  geom_point() +
  geom_errorbar(aes(ymin = CI_lower, ymax = CI_upper)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

<br />

## Problem 2

```{r load_clean_data_2}
birth_weight_df =
  read_csv("./birth_weight_data/birthweight.csv") %>% 
  janitor::clean_names() %>% 
  drop_na() %>% 
  filter(frace != 9) %>% 
  mutate(
    babysex = case_when(
      babysex == 1 ~ "male",
      babysex == 2 ~ "female"
    ),
    malform = case_when(
      malform == 0 ~ "absent",
      malform == 1 ~ "present"
    ),
    frace = case_when(
      frace == 1 ~ "White",
      frace == 2 ~ "Black",
      frace == 3 ~ "Asian",
      frace == 4 ~ "Puerto Rican",
      frace == 8 ~ "Other"
    ),
    mrace = case_when(
      mrace == 1 ~ "White",
      mrace == 2 ~ "Black",
      mrace == 3 ~ "Asian",
      mrace == 4 ~ "Puerto Rican",
      mrace == 8 ~ "Other"
    )
  ) %>% 
  mutate(
    across(where(is.character), as.factor)
  )
```

<br />

### Build models - Step-wise Backward Method

Model 1: 

The assumption is that mother's physical characteristics have effects on baby birth weights. 

First, fit the model for all predictors.

```{r}
fit_df =
  birth_weight_df %>% 
  select(bwt, delwt, gaweeks, menarche, mheight, momage, ppbmi, ppwt, smoken)

mul_fit = lm(bwt ~ ., data = fit_df)

summary(mul_fit)
```

Use function step() for variable selection based on AIC criterion, the option used is "backward".

```{r}
step(mul_fit, direction = "backward")
```

The goal is to have the lowest AIC for all predictors combined. The output showed that AIC increased after taking out each variable, thus, all variables are significant predictors for baby birth weights. We should keep all predictors.

```{r}
model_1 = lm(bwt ~ ., data = fit_df)

model_1 %>% 
  broom::tidy() %>% 
  select(-std.error, -statistic) %>% 
  knitr::kable()
```

Look at residuals

```{r}
fit_df =
  fit_df %>% 
  modelr::add_residuals(model_1) %>% 
  modelr::add_predictions(model_1)

fit_df
```

Make residual plots

```{r}
fit_df %>% 
  ggplot(aes(x = pred, y = resid)) +
  geom_point(alpha = 0.5)
```



```{r}
fit_df %>% 
  ggplot(aes(x = resid)) +
  geom_density()
```


<br />

### Other Two Models

Model 2: Length at birth and gestational age as predictors

```{r}
model_2 = lm(bwt ~ blength + gaweeks, data = birth_weight_df)

model_2 %>% 
  broom::tidy() %>% 
  select(-std.error, -statistic) %>% 
  knitr::kable()
```

Model 3: Head circumference, length, sex and all interactions between these.

```{r}
model_3 = lm(bwt ~ bhead * blength * babysex, data = birth_weight_df)

model_3 %>% 
  broom::tidy() %>% 
  select(-std.error, -statistic) %>% 
  knitr::kable()
```

<br />

## Problem 3

```{r load_clean_data_3}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

